---
title: "Geometric Ornstein-Uhlenbeck process"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geometric Ornstein-Uhlenbeck process}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
```

This document describes the fundamentals of the geometric Ornstein-Uhlenbeck (OU) process and its implementation in the **tsmodels** package.

The geometric Ornstein-Uhlenbeck (OU) process, like its classical version, models mean-reverting stochastic processes. It is stationary, Gaussian and Markov. Mathematically, we can define the geometric OU process via the Langevin stochastic differential equation
$$dY_t = - \frac{(Y_t-\mu)}{\tau}dt + \sigma Y_t dW_t.$$

Here, because of the geometric nature of the stochastic term, $\mu$ and $\sigma$ are not directly the process mean and standard deviation, although they are related. As usual $W$ is a Brownian motion (or Wiener process).  

While the paths in the classic OU process can be negative in value, in the geometric version the process is always positive semi-definite. This means that we can model a mean-reverting price time series via the geometric version of the Ornstein-Ulhenbeck process. 

Here, we integrate the above stochastic differential equation using the Euler-Maruyama method, which is described in more details in the vignette describing the classic Ornstein-Uhlenbeck process. 

Let's see how to generate realizations of the geometric OU process:


```{r, echo=TRUE}
library(tsmodels)

T <- 100
dt <- 0.01
mu <- 5.0
sigma <- 1.0
tau <- 5
Y0 <- 10

result <- geometric_Ornstein_Uhlenbeck(T=T, dt=dt, mu=mu, sigma=sigma, tau=tau, Y0=Y0)
t <- result[[1]]
Y <- result[[2]]
```

Let's plot the results. We need to load the **tidyverse** package to use ggplot2

```{r echo=TRUE, message=FALSE}
library(tidyverse)
```

```{r, echo=TRUE}
ggplot()+
    geom_line(mapping = aes(x=t, y=Y), color="black")
```
